graphsim: An R package for simulating gene expression data from graph structures of biological pathways

### 2020-04-30 paper/paper.Rmd
Rmarkdown figure 2 #13

```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.455\\linewidth', out.width='.455\\linewidth', fig.retina=10, fig.margin = FALSE, fig.cap = '\\textbf{Simulated graph structures.}. A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3. Activating links are denoted by black arrows and inhibiting links by red edges. Inhibiting edges have been highlighted in red.'}
#, fig.subcap=c("Activating pathway structure", "Inhibiting pathway structure")}
#generate graph structure

@@

![\small \textbf{\textbf{Simulating expression from a graph structure.}}
An example of a graph structure (a) that has been used to derive a
relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
(d) from the relative distances between the nodes. Non-negative values
are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
used to generate a simulated expression dataset of 100 samples (coloured
blue to red from low to high) via sampling from the multivariate
normal distribution. Here genes with closer relationships in the pathway
structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating label="simulation_activating"}

@@

```r
# activating graph
state <- rep(length(E(graph)), 1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2)

#adjacency matrix

adj_mat <- make_adjmatrix_graph(graph)

#relationship matrix

dist_mat <- make_distance_graph(graph, absolute = FALSE)

#sigma matrix directly from graph

sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)

#show shortest paths of graph

shortest_paths <- shortest.paths(graph)

#generate expression data directly from graph

expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
          dist = TRUE, absolute = FALSE, state = state)

#plot adjacency matrix

library("gplots")

heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot relationship matrix

heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot sigma matrix

heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

@@

 \newpage

```r
expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
          comm = FALSE, dist =TRUE, absolute = FALSE, state = state)

#plot simulated expression data

heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot simulated correlations

heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

>>>>

```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}. A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3. Activating links are denoted by black arrows and inhibiting links by red edges. Inhibiting edges have been highlighted in red.'}
#generate graph structure

@@

@@

```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}. An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
# activating graph
state <- rep(1, length(E(graph)))
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
              sub = "(a) Activating pathway structure", cex.sub = 1.5)

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)
#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)
#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
#show shortest paths of graph
shortest_paths <- shortest.paths(graph)
#generate expression data directly from graph
expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE, state = state)
#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.5)
#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.5)
#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expressino data (log scale)", cex.sub = 1.5)
#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

%![\small \textbf{\textbf{Simulating expression from a graph structure.}}
%An example of a graph structure (a) that has been used to derive a
%relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
%(d) from the relative distances between the nodes. Non-negative values
%are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
%used to generate a simulated expression dataset of 100 samples (coloured
%blue to red from low to high) via sampling from the multivariate
%normal distribution. Here genes with closer relationships in the pathway
%structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating %label="simulation_activating"}

@@
